### 2020-01-24 defesa/tex/chapter/mcd.tex
wip

\begin{equation}
\max_{Q} w\left(\bigcup_{j=1}^m \Pp \cap D_j(q_j)\right).
\end{equation}

@@

This version of the problem will be refered to as \sigla{MCD-1}{Maximum Covering by One Unit Disk} and it is just a specific case of MCD with only one disk with radius one ($m=1$ and $r_1=1$). We later use the algorithm for MCD-1 here described to construct a CLS which is guaranteed to contain an optimal solution for MCD.

Two exact methods for MCD-1 have been found in the literature. A $\bigO(n^2)$ algorithm is proposed by \citeonline{chazelle:1986} which improved the previously $\bigO(n^2\log{n})$ one proposed by \citeonline{drezner}.
As it has been mentioned, MCD-1 is a 3SUM-HARD problem, which means that it is as hard as the 3SUM problem (the problem of finding three real numbers that sum to zero, given $n$ real numbers). Initially the lower bound of the 3SUM problem was conjectured to be $\Omega(n^2)$, matching the best algorithm for MCD-1, which meant that no better time-complexity could be achieved. Since then, however, better algorithms for 3SUM have been developed with a $\bigO(\frac{n^2}{poly(n)})$ run time complexity \cite{3SUM-kopelowitz:2014}.

@@

In \citeonline{chazelle:1986} and \citeonline{cabello:2006}, on the other hand, instead of working directly with MCD-1, an equivalent problem called \sigla{MWC}{Maximum Weight Clique} was introduced. The algorithm for MCD that is developed in this chapter also uses that equivalence. Because of that, MWC is introduced in the next section.

@@

\begin{equation}
S = \bigcap_{j : q \in D_j} D_j.
\end{equation}
With a geometric observation, though, the number of possible values for the solution can be reduced.
Let $\partial \D = \{ \partial D_1, \dots, \partial D_n \}$ be the set of circles corresponding to each disk in $\D$, it is possible to see that any clique has at least one point on its border that is in the intersection of two circles in $\partial D$ or, in the case where the clique is composed of only one disk, it contains the center of its only disk. Because of that, $q$ can be actually limited to the set of pairwise intersections of $\partial \D$ as well as the set of centers of each disk $\Pp$. Then, with this observation, an optimal solution of MWC can be defined by the optimization problem

\begin{equation}
\max_{q} \sum_{D_k \cap q \neq \emptyset} w_k,
\end{equation}
with $q \in \{\partial D_i \cap \partial D_j : 1 \le i < j \le n\} \cup \Pp$. With this new specification of the solution's search space, it can be shown that given an instance of MWC, an optimal solution for it can be found by going through $\binom{n}{2} + n = \bigO(n^2)$ possible ones.

@@

As it has been mentioned, with the equivalence of the two problems, an optimal solution of the Maximum Weight Clique Problem is also an optimal solution of MCD-1, which means that a disk centered at $q^*$ which is an optimal solution of MWC, will have a maximal weight covering of the demand set $\Pp$.

Given an instance of MCD-1, the equivalent MWC intance is obtained by defining the set $\D$ to contain the disks centered at $\Pp$ and setting the weight of every disk to be the weight of its corresponding point in $\Pp$. A disk $D_i$ will represent the area where a disk can be placed in order to cover $p_i$. This means that a intersection between some disks is a region where a disk could be placed to cover the corresponding points.

@@

Formally, in MWC, if a point $q$ lies inside $\bigcap_{k \in I} D_k$, with $I \subset \{1,\dots,n\}$, then a disk centered at $q$ will cover the points $p_k$, with $k\in I$ in the equivalent MCD-1 instance. Conversely, the same applies for a disk placed at $q$ that covers points $p_k$, with $k \in I$ in the MCD-1 instance. It means that $q$ will lie inside region $\bigcap_{k \in I} D_k$ in MWC.

@@

\begin{definicao}\label{def:inter_arc}
    Let $D_i$ and $D_j$ be two unit disks with non-empty intersection, and $(\theta_1, \theta_2) \in [0,2\pi)^2$ be the two angles that $\partial D_i$ and $\partial D_j$ intersect, with the condition that $(\theta_1,\theta_2)$ defines an arc (counter-clockwise order) of $D_i$ that is the border of $D_i \cap D_j$. Then, define $\Gamma_+(i,j) = \theta_1$ and $\Gamma_-(i,j) = \theta_2$. For convenience, if $D_i$ is tangent to $D_j$, then $\theta_1=\theta_2$; and if $i=j$, then $\Gamma_-(i, j)=0$ and $\Gamma_+(i,j)=2\pi$.
\end{definicao}

@@

With \autoref{def:inter_arc} in hand, we can establish the basis of the algorithm for MWC. For every disk $D_i$, let us describe an algorithm that gets the best clique which $D_i$ is part of. This way, an algorithm for MWC just uses that method for every disk and returns the best solution found. Firstly, let $A_i$ be a list that contains the intersection angles of $\partial D_i$ with every circle in $\partial \D$. Assume also that $A_i$ is actually a circular list sorted in ascending order and it is defined as

\begin{equation}
A_i = \bigcup_{j=1}^n \{ \Gamma_-(i,j), \Gamma_+(i,j) \}.
\end{equation}
Finding the best solution which $D_i$ is part of can be done by traversing $A_i$ while keeping a set of active disks. When an opening intersection angle is reached, the corresponding disk is added to the active set; and when a closing one is seen, the corresponding disk is removed from the active set. This way, finding an optimal solution can be achieved by keeping the weight of the active disks as well as the best clique found so far. Notice also that because $\Gamma_+(i,i)=0$ and $\Gamma_-(i,i)=2\pi$, any clique found by the traversal will also contain $D_i$.

In practice, traversing a circular list can be emulated by traversing a regular list that has a copy of the original circular list added to its end. This is done here by traversing the regular list $B_i$ defined as

\begin{equation}\label{eq:b_i}
B_i = A_i\cup\bigcup_{j=1}^n \{ 2\pi+\Gamma_-(i,j), 2\pi+\Gamma_+(i,j) \}.
\end{equation}
Assuming $B_i$ is sorted in ascending order, a copy of the elements of $A_i$ shifted to the interval $[2\pi, 4\pi]$ is added to its end and a simple traversal, going through the first element until the last one, simulates a traversal on the circular list $A_i$.
This works because for any pair of disks, such that $\Gamma_+(i,j) > \Gamma_-(i,j)$, $B_i$ will contain $\Gamma_-(i,j) < \Gamma_+(i,j) + 2\pi$. This is shown in \autoref{fig:array_disks}, where the list of intersection between $\partial D_1$ with circles $\partial D_2, \partial D_3$ and $\partial D_4$ is duplicated simulating a traversal in $B_i$ (note the indication to where the traversal starts as well as the positive and negative signs representing when a intersection with another disk opens and closes, respectively).

@@

    \caption{The intersections list of a disk with three other disks.}
	 %\input{tex/figures/array_disks.tex}
    \includegraphics[scale=.3]{tex/figures/3_disks_intersect2.pdf}
    \fautor
    \label{fig:array_disks}
\end{figure}

@@

\begin{algoritmo}[b]
\caption{Algorithm for MCD-1.}\label{algoritmo:mcd_1}
\begin{algorithmic}[1]
\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$ with weights $\Ww=\{w_1, \dots, w_n\}$.}
\Ensure{The maximum cover by a unit disk.}

@@

\Procedure{$MCD_1$}{$\Pp$}
\State $Q_{best} \gets \{\}$
\State $ans \gets$ center of $D_1$
\ForAll{$p_i \in \Pp$}
\State Let $D_i$ be the disk with center at $p_i$
\State Let $B_i$ be the list of intersection angles of $D_i$ as defined by \autoref{eq:b_i}.

\State $Q \gets \{\}$ \Comment{The set of active disks.}
\For{$a \in B_i$}\Comment{Assuming $B_i$ is sorted.}
\State Let $D_a$ be the disk that intersects $D_i$ at angle $a$. 
\If{$a$ is a starting angle}
\State $Q \gets Q \cup \{D_a\}$
\Else
\State $Q \gets Q \setminus \{D_a\}$
\EndIf
\If{$w(Q_{best}) < w(Q)$} \Comment{$w(Q)$ returns the weights of the disks in $Q$.}
\State $Q_{best} \gets Q$
\State $ans \gets $ point corresponding to the intersection angle $a$
\EndIf

@@

\State \Return $ans$
\EndProcedure
\end{algorithmic}
\end{algoritmo}

@@

A simple adaptation can be done on \autoref{algoritmo:mcd_1} to make it return a CLS that contains an optimal solution of MCD for that disk. This is shown in \autoref{algoritmo:mcd_cls}. 
Notice also that MCD-1 is defined only for unit disks, however, this constraint can be dropped, as it is introduced just for the sake of keeping the text more simple and \autoref{algoritmo:mcd_1} works for any radius. A result about the runtime complexity of \autoref{algoritmo:mcd_cls} has already been given by \autoref{lema:disk}, the following result states about the adaption of it to be used in an algorithm for MCD.

@@

\begin{algoritmo}
	\caption{Algorithm for MCD-1 that returns a CLS.}\label{algoritmo:mcd_cls}
	\begin{algorithmic}[1]
		\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$ with weights $\Ww=\{w_1, \dots, w_n\}$, and a radius $r\in\R_{>0}$.}
		\Ensure{A CLS for the disk given by radius $r$.}

@@

		\Procedure{$MCD_1$}{$\Pp, \Ww, r$}
		\State $S \gets \{\}$
		\ForAll{$p_i \in \Pp$}
		\State Let $D_i$ be the disk with center at $p_i$
		\State Let $B_i$ be the list of intersection angles of $D_i$ as defined by \autoref{eq:b_i}.

		\State $Cov \gets \{\}$
		\For{$a \in B_i$}\Comment{Assuming $B_i$ is sorted.}
		\State Let $q_a$ be the intersection point correspondent to angle $a$. 
		\State $S \gets S \cup \{q_a\}$	
		\EndFor
		\EndFor

>>>>

\begin{equation*}
\max_{Q} w\left(\bigcup_{j=1}^m \Pp \cap D_j(q_j)\right).
\end{equation*}

@@

This version of the problem will be refered to as \sigla{MCD1}{Maximum Covering by One Unit Disk} and it is just a specific case of MCD with only one disk with radius one ($m=1$ and $r_1=1$). We refer to an instance of MCD1 as the tuple $(\Pp, \Ww)$.
We later use the algorithm for MCD1 here described to construct a CLS which is guaranteed to contain an optimal solution for MCD.

Two exact methods for MCD1 have been found in the literature. A $\bigO(n^2)$ algorithm is proposed by \citeonline{chazelle:1986} which improved the previously $\bigO(n^2\log{n})$ one proposed by \citeonline{drezner}.
As it has been mentioned, MCD1 is a 3SUM-HARD problem, which means that it is as hard as the 3SUM problem (the problem of finding three real numbers that sum to zero, given $n$ real numbers). Initially the lower bound of the 3SUM problem was conjectured to be $\Omega(n^2)$, matching the best algorithm for MCD1, which meant that no better time-complexity could be achieved. Since then, however, better algorithms for 3SUM have been developed with a $\bigO(\frac{n^2}{poly(n)})$ run time complexity \cite{3SUM-kopelowitz:2014}.

@@

In \citeonline{chazelle:1986} and \citeonline{cabello:2006}, on the other hand, instead of working directly with MCD1, an equivalent problem called \sigla{MWC}{Maximum Weight Clique} was introduced. The algorithm for MCD that is developed in this chapter also uses that equivalence. Because of that, MWC is introduced in the next section.

@@

\begin{equation*}
S = \bigcap_{j : q \in D_j} D_j.
\end{equation*}
With a geometric observation, though, the number of possible values for the solution can be reduced.
Let $\partial \D = \{ \partial D_1, \dots, \partial D_n \}$ be the set of circles corresponding to each disk in $\D$, it is possible to see that any clique has at least one point on its border that is in the intersection of two circles in $\partial D$ or, in the case where the clique is composed of only one disk, it contains the center of its only disk. Because of that, $q$ can be actually limited to the set of pairwise intersections of $\partial \D$ as well as the set of centers of each disk $\Pp$. Then, with this observation, an optimal solution of MWC can be defined by the optimization problem

\begin{equation*}
\max_{q} \sum_{D_k \cap q \neq \emptyset} w_k,
\end{equation*}
with $q \in \{\partial D_i \cap \partial D_j : 1 \le i < j \le n\} \cup \Pp$. With this new specification of the solution's search space, it can be shown that given an instance of MWC, an optimal solution for it can be found by going through $\binom{n}{2} + n = \bigO(n^2)$ possible ones.

@@

As it has been mentioned, with the equivalence of the two problems, an optimal solution of the Maximum Weight Clique Problem is also an optimal solution of MCD1, which means that a disk centered at $q^*$ which is an optimal solution of MWC, will have a maximal weight covering of the demand set $\Pp$.

Given an instance of MCD1, the equivalent MWC intance is obtained by defining the set $\D$ to contain the disks centered at $\Pp$ and setting the weight of every disk to be the weight of its corresponding point in $\Pp$. A disk $D_i$ will represent the area where a disk can be placed in order to cover $p_i$. This means that a intersection between some disks is a region where a disk could be placed to cover the corresponding points.

@@

Formally, in MWC, if a point $q$ lies inside $\bigcap_{k \in I} D_k$, with $I \subset \{1,\dots,n\}$, then a disk centered at $q$ will cover the points $p_k$, with $k\in I$ in the equivalent MCD1 instance. Conversely, the same applies for a disk placed at $q$ that covers points $p_k$, with $k \in I$ in the MCD1 instance. It means that $q$ will lie inside region $\bigcap_{k \in I} D_k$ in MWC.

@@

\begin{definicao}\label{def:inter_arc}
    Let $D_i$ and $D_j$ be two unit disks with non-empty intersection, and $(\theta_1, \theta_2) \in [0,2\pi)^2$ be the two angles that $\partial D_i$ and $\partial D_j$ intersect, with the condition that $(\theta_1,\theta_2)$ defines an arc (counter-clockwise order) of $D_i$ that is the border of $D_i \cap D_j$. Then, define $\Gamma_+(i,j) = \theta_1$ and $\Gamma_-(i,j) = \theta_2$. For convenience, if $D_i$ is tangent to $D_j$, then $\theta_1=\theta_2$; and if $i=j$, then $\Gamma_+(i, j)=0$ and $\Gamma_-(i,j)=2\pi$.
\end{definicao}

@@

With \autoref{def:inter_arc} in hand, we can establish the basis of the algorithm for MWC. For every disk $D_i$, let us describe an algorithm that gets the best clique which $D_i$ is part of. This way, an algorithm for MWC just uses that method for every disk and returns the best solution found. Firstly, let $A_i$ be a circular list that contains the intersection angles of $\partial D_i$ with every circle in $\partial \D$ defined as

\begin{equation*}
A_i = \bigcup_{j=1}^n \{ \Gamma_-(i,j), \Gamma_+(i,j) \}.
\end{equation*}
Assume also that $A_i$ is sorted in ascending order by the angle values with ties being broken by prioritizing opening angles.

Finding the best solution which $D_i$ is part of can be done by traversing $A_i$ while keeping a set of active disks. When an opening intersection angle is reached, the corresponding disk is added to the active set; and when a closing one is seen, the corresponding disk is removed from the active set. This way, finding an optimal solution can be achieved by keeping the weight of the active disks as well as the best clique found so far. Notice also that because $\Gamma_+(i,i)=0$ and $\Gamma_-(i,i)=2\pi$, any clique found by the traversal will also contain $D_i$.

In practice, traversing a circular list can be emulated by traversing a regular list that has a copy of the original circular list added to its end. 
Therefore, the list $B_i$ is defined here as a list that contains the elements of $A_i$ and a copy of it shifted to the interval $[2\pi, 4\pi]$. It is defined as

\begin{equation}\label{eq:b_i}
B_i = A_i\cup\bigcup_{j=1}^n \{ 2\pi+\Gamma_-(i,j), 2\pi+\Gamma_+(i,j) \}.
\end{equation}
Assuming $B_i$ is sorted with the same criteria as $A_i$. A simple traversal, starting at the first element and going until the last one, simulates a traversal on the circular list $A_i$.
This works because for any pair of disks $D_i$, $D_j$; $B_i$ contains $\Gamma_+(i,j) < \Gamma_-(i,j) + 2\pi$. That is, the algorithm encounters an opening angle before reaching a closing one for any circle.

In \autoref{fig:array_disks}, the intersection points between $\partial D_1$ (solid border) with $\partial D_2, \partial D_3$, and $\partial D_4$ (dashed border) are shown with a plus or minus sign indicating opening or closing intersection angles. 
The intersection list $B_1$ is also displayed in \autoref{fig:array_disks} along with the size of the set of active disks $Q$ after processing every point in $B_1$ -- this is exactly what \autoref{algoritmo:mcd_1} for MWC does for every disk. The region of optimal solutions is highlighted and can also be obtained by following $B_1$ while keeping track of $Q$.
The special intersection point of $\partial D_1$ with itself can also be seen in \autoref{fig:array_disks}. Its usage is very convenient as with $\Gamma_+(1, 1)$ and $\Gamma_-(1,1)$  in $B_1$, the algorithm inserts $D_1$ in the set of active disks before processing any point and removes $D_1$ only after every point has been processed.

Finally, we define \autoref{algoritmo:mcd_1} for MWC. Given an instance $(\Pp, \Ww)$ of MWC, the algorithm uses the approach described here of keeping a set of active disks while traversing the list $B_i$. It returns a point that is inside an optimal clique, this way \autoref{algoritmo:mcd_1} can also be used to get an optimal solution for an instance $(\Pp, \Ww)$ of MCD1.

@@

    \caption{The intersection list of a disk with three other disks.}
	 %\input{tex/figures/array_disks.tex}
    \includegraphics[scale=.3]{tex/figures/3_disks_intersect2.pdf}
    \fautor
    \label{fig:array_disks}
\end{figure}

@@

\begin{algoritmo}[b]
\caption{Algorithm for MWC.}\label{algoritmo:mcd_1}
\begin{algorithmic}[1]
\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, and a set of weights $\Ww=\{w_1, \dots, w_n\}$.}
\Ensure{A point that is inside the maximum weight clique of unit disks.}

@@

\Procedure{$MWC$}{$\Pp, \D, \Ww$}
\State Let $\D=\{D_1(p_1), \dots, D_n(p_n)\}$ be a set of unit disks
\State $Q_{best} \gets \{\}$
\State $q^* \gets$ $p_1$
\ForAll{$D_i \in \Pp$}
\State Let $B_i$ be the list of intersection angles of $D_i(p_i)$ as defined by \autoref{eq:b_i}

\State $Q \gets \{\}$ \Comment{The set of active disks.}
\For{$\theta \in B_i$}\Comment{Assuming $B_i$ is sorted.}
\State Let $D_j$ be the disk, such that $\theta \in D_j(p_j)\cap D_i(p_i)$
\If{$\theta$ is a opening angle}
\State $Q \gets Q \cup \{D_j\}$
\Else
\State $Q \gets Q \setminus \{D_j\}$
\EndIf
\If{$w(Q_{best}) < w(Q)$} 
\State $Q_{best} \gets Q$
\State $q^* \gets$ point corresponding to the intersection angle $\theta$
\EndIf

@@

\State \Return $q^*$
\EndProcedure
\end{algorithmic}
\end{algoritmo}

@@

A simple adaptation can be done on \autoref{algoritmo:mcd_1} to make it return a CLS that contains an optimal solution of MCD for that disk. This is shown in \autoref{algoritmo:mcd_cls}. 
Notice also that MCD1 is defined only for unit disks, however, this constraint can be dropped, as it is introduced just for the sake of keeping the text more simple and \autoref{algoritmo:mcd_1} works for any radius. A result about the runtime complexity of \autoref{algoritmo:mcd_cls} has already been given by \autoref{lema:disk}, the following result states about the adaption of it to be used in an algorithm for MCD.

@@

\begin{algoritmo}
	\caption{Algorithm for MCD1 that returns a CLS.}\label{algoritmo:mcd_cls}
	\begin{algorithmic}[1]
		\Require{A set of points $\Pp=\{p_1,\dots,p_n\}$ with weights $\Ww=\{w_1, \dots, w_n\}$, and a radius $r\in\R_{>0}$.}
		\Ensure{A CLS for the disk given by radius $r$.}

@@

		\Procedure{CLS-MCD}{$\Pp, \Ww, r$}
		\State $S \gets \{\}$
		\ForAll{$p_i \in \Pp$}
		\State Let $D_i(p_i)$ be the disk with center at $p_i$
		\State Let $B_i$ be the list of intersection angles of $D_i$ as defined by \autoref{eq:b_i}

		\For{$\theta \in B_i$}\Comment{Assuming $B_i$ is sorted.}
		\If{$\theta$ is a opening angle}
		\State Let $q_\theta$ be the intersection point correspondent to angle $\theta$
		\State $S \gets S \cup \{q_\theta\}$	
		\EndIf
		\EndFor
		\EndFor

### 2020-01-24 defesa/tex/chapter/mce.tex
wip

\begin{equation}
\max_q w\left(\bigcup_{j=1}^m \Pp \cap E_j(q_j)\right).
\end{equation}

@@

\begin{equation}
\max_q w(\Pp \cap E(q)).
\end{equation}
An adaptation of \autoref{algoritmo:mcd_1} is obtained by just replacing the function that finds the intersection points between two disks by a function that finds the intersection points between two ellipses $\partial E_i$ and $\partial E_j$.
It can be seen in \autoref{fig:3ellipses_intersect} that the intersection points and their correspondents $\Gamma_-(i,j)$ and $\Gamma_+(i,j)$ functions behave the same way as in the disks case.
The intersection of two ellipses as well as determining $\Gamma_-(i,j)$ and $\Gamma_+(i,j)$ are described thoroughly in \autoref{chapter:ellipses_intersection}.

\begin{figure}[H]
\centering
    \caption{Three ellipses and their intersection points}
    \includegraphics[scale=.32]{tex/figures/3ellipses_intersect.pdf}
    \fautor
    \label{fig:3ellipses_intersect}
\end{figure}
\section{An algorithm for MCE}

The same procedure defined in \autoref{algoritmo:mcd_cls} can be used to get a CLS for every ellipse in MCE. We refer to the elliptical version of that procedure as $MCE_1$ (we do not define it in this chapter because it would look the same as $MCD_1$, apart from the name, of course).

@@

It is worth mentioning that, even though we call $MCE_1$ every time in the recursion, in practice, it is probably best to pre-process this step, and only call it $m$ times for the whole set of points.
Some other easy improvements can also be made in the implementation. For example, if an ellipse covers two sets of points $X$ and $Y$, with $X \subset Y$, then set $X$ can be ignored by the algorithm because of the non-negative weights constraint. Also, if two ellipses have their centers with Euclidean distance greater than their semi-major parameter, they for sure do not intersect. Depending on the input, this observation can make the algorithm not go through the whole list of ellipses every time it needs to determine the ellipses pairwise intersections.

@@

We refer to this version of the problem as  \sigla{MCE-k}{Maximum Covering by Ellipses with a $k$-constraint}. An instance of it is given by the same parameters as MCE, plus a list of costs $\Cc=\{c_1, \dots, c_m\}$, with $c_j\in\R_{\ge0}$ being the $j$-th ellipse's cost, and $k\in\mathbb{N}$.

 A solution for MCE-k, however,  when compared to MCE's, has a bit more cluttered description. We define it as a set $I:=\{i_1, \dots, i_k\}\subset\{1, \dots, m\}$, such that $|I|=k$; and a tuple $Q:=(q_1, \dots, q_k)$, with $q_j\in\R^2$ being the center of the $j$-th ellipse in $I$. An optimal solution of MCE-k is given by the optimization problem

\begin{equation}
\max_{I, Q} w\left(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j)\right).
\end{equation}

Finally, \autoref{algoritmo:mce} can serve as basis for an algorithm for MCE-k. 
Firstly, for every subset $I \subset \{1, \dots, m\}$, such that $|I| = k$, the algorithm for MCE is invoked for the instance $(\Pp, \Ww, \{(a_j, b_j) : j \in I\})$; that is, an instance where only the ellipses in $I$ are present.
After that, by keeping track of the utilized ellipses' costs for every $I \subset \{1, \dots, m\}$, an optimal solution can be obtained.
This simple adjustment achieves a run-time complexity of $\bigO(\binom{m}{k} \times n^{2k})$.

\begin{algoritmo}
    \caption{Algorithm for MCE}\label{algoritmo:mce}
    \begin{algorithmic}[1]
        \Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, a list of weights $\Ww=\{w_1, \dots, w_n\}$, and a list of shape parameters $\Rr=\{(a_1, b_1), \dots, (a_m, b_m)\}$.}

        \Ensure{The value of an optimal solution of MCE.}

@@

        \Procedure{$MCE$}{$\Pp, \Ww, \Rr, j=1$}
        \If{$j = m+1$}
        \State \Return $0$
        \EndIf

        \State $ans \gets 0$

        \State $S_j \gets MCE_1(\Pp, a_j, b_j)$
        \For{$q \in S_j$}
        \State $Cov \gets \Pp \cap q$
        \State $ans \gets \max\{ans, w(Cov) + MCE(\Pp \setminus Cov, \Ww, \Rr, j+1)\}$ \Comment{Calls the procedure for the next ellipse.}
        \EndFor

        \State \Return $ans$
        \EndProcedure
    \end{algorithmic}
\end{algoritmo}

>>>>

\begin{equation*}
\max_q w\left(\bigcup_{j=1}^m \Pp \cap E_j(q_j)\right).
\end{equation*}

@@

\begin{equation*}
\max_q w(\Pp \cap E(q)).
\end{equation*}
An adaptation of \autoref{algoritmo:mcd_1} is obtained by just replacing the function that finds the intersection points between two disks by a function that finds the intersection points between two ellipses $\partial E_i$ and $\partial E_j$.
It can be seen in \autoref{fig:3ellipses_intersect} that the intersection points and their correspondents $\Gamma_-(i,j)$ and $\Gamma_+(i,j)$ functions behave the same way as in the disks case.
The intersection of two ellipses as well as determining $\Gamma_-(i,j)$ and $\Gamma_+(i,j)$ are described in the next section.

\begin{figure}[H]
	\centering
	\caption{Three ellipses and their intersection points}
	\includegraphics[scale=.34]{tex/figures/3ellipses_intersect.pdf}
	\fautor
	\label{fig:3ellipses_intersect}
\end{figure}

\section{Determining $\Gamma_+(i,j)$ and $\Gamma_-(i,j)$}\label{section:ellipses_intersection}

Let $E_1(q_1)$, and $E_2(q_2)$ be two coverage region of ellipses centered at $q_1, q_2\in\R^2$ respectively, with shape parameters $(a, b)\in\R^2_{>0}$. After changing the coordinates to make the center of the first ellipse be at the origin, the intersection points between the two ellipses are defined by

\begin{align}
\frac{x^2}{a^2} + \frac{y^2}{b^2} = 1 && (E_1) \label{eq:ell_inter_1}\\
\frac{(x-h)^2}{a^2} + \frac{(y-k)^2}{b^2} = 1 && (E_2), \nonumber
\end{align}
where $(h,k)\in\R^2$ is the center of the second ellipse after the coordinates were translated by $-q_1$. As both equations are equal to $1$, we have
\begin{align}
    b^2x^2 + a^2y^2 &= b^2(x-h)^2 + a^2(y-k)^2 \nonumber\\
    x &= y\frac{-2ka^2}{2hb^2} + \frac{b^2h^2 + a^2k^2}{2hb^2} \nonumber\\
    x &= y\alpha + \beta.\label{eq:ell_inter_2}
\end{align}
Replacing \autoref{eq:ell_inter_2} into \autoref{eq:ell_inter_1}, we get
\begin{align}\label{eq:ell_inter_3}
y^2(b^2\alpha^2 + a^2) + y(2\beta\alpha b^2) + b^2\beta^2 -a^2b^2 = 0,
\end{align}
Which is a second degree polynomial. Then, $\partial E_1(q_1) \cap \partial E_2(q_2) \neq \{\}$ if, and only if the roots of \autoref{eq:ell_inter_3} are real. The intersection points itself can be obtained by solving the polynomial for $y$ and applying its value onto the $x=y\alpha + \beta$ equation.

Suppose that $\partial E_1(q_1) \cap \partial E_2(q_2) = \{p_1, p_2\}$, with $p_1 \neq p_2$. To determine $\Gamma_+(1,2)$ and $\Gamma_-(1,2)$, we need to first determine the angles of intersection of $p_1$ and $p_2$ on $E_1(q_1)$ using the parametric equation for an ellipse defined in \autoref{eq:parametric_ellipse}. Given a point $(x, y)\in \partial E_1(q_1)$, using \autoref{eq:parametric_ellipse} we have

\begin{equation*}
\dfrac{a}{b}\dfrac{y-q_{1y}}{x-q_{1x}} = \tan{t}.
\end{equation*}
Solving for $t$ gives the angle of point $(x,y)$ on $E_1(q_1)$. Using that process we obtain $t^{(1)}_1, t^{(1)}_2$ as the angles of the points $\{p_1, p_2\}$ on $E_1(q_1)$; and $t^{(2)}_1, t^{(2)}_2$ as the angles of points $\{p_1, p_2\}$ on $E_2(q_2)$.
Let $\gamma_1$ be the curve defined by $E_1(q_1)$ and $\gamma_2$ the curve defined by $E_2(q_2)$. Considering its derivatives \autoref{eq:der_parametric_ellipse} at the intersection points, as it can be seen in \autoref{fig:ellipse_gamma}, we have that $p_1$ is $\Gamma_+(1,2)$ if, and only if
\begin{equation*}
angle(\gamma_1'(t_1^{(1)}), -\gamma_2'(t_1^{(2)}) < \pi.
\end{equation*}
The case where that angle is equal to $\pi$ happens when both ellipses intersect at only one point. This case has to be treated separately as we need to have two equal intersection points: one as $\Gamma_+(1,2)$ and the other as $\Gamma_-(1,2)$.

\begin{figure}
	\centering

	\caption{The intersection list of a disk with three other disks.}
	%\input{tex/figures/array_disks.tex}
	\includegraphics[scale=.38]{tex/figures/ellipse_gamma.pdf}
	\fautor
	\label{fig:ellipse_gamma}
\end{figure}

\section{An algorithm for MCE}

The same procedure defined in \autoref{algoritmo:mcd_cls} can be used to get a CLS for every ellipse in MCE. We refer to the elliptical version of that procedure as CLS-MCE (we do not define it in this chapter because it would look the same as CLS-MCD defined in \autoref{algoritmo:mcd_cls}, apart from the name, of course).

@@

It is worth mentioning that, even though we call CLS-MCE every time in the recursion, in practice, it is probably best to pre-process this step, and only call it $m$ times for the whole set of points.
Some other easy improvements can also be made in the implementation. For example, if an ellipse covers two sets of points $X$ and $Y$, with $X \subset Y$, then set $X$ can be ignored by the algorithm because of the non-negative weights constraint. Also, if two ellipses have their centers with Euclidean distance greater than their semi-major parameter, they for sure do not intersect. Depending on the input, this observation can make the algorithm not go through the whole list of ellipses every time it needs to determine the ellipses pairwise intersections.

@@

We refer to this version of the problem as  \sigla{MCE-$k$}{Maximum Covering by Ellipses with a $k$-constraint}. An instance of it is given by the same parameters as MCE, plus a list of costs $\Cc=\{c_1, \dots, c_m\}$, with $c_j\in\R_{\ge0}$ being the $j$-th ellipse's cost, and $k\in\mathbb{N}$, $k\le m$.

 A solution for MCE-$k$, however,  when compared to MCE's, has a bit more cluttered description. We define it as a set $I:=\{i_1, \dots, i_k\}\subset\{1, \dots, m\}$, such that $|I|=k$; and a tuple $Q:=(q_1, \dots, q_k)$, with $q_j\in\R^2$ being the center of the $j$-th ellipse in $I$. An optimal solution of MCE-$k$ is given by the optimization problem

\begin{equation*}
\max_{I, Q} w\left(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j)\right).
\end{equation*}

Finally, \autoref{algoritmo:mce} can serve as basis for the \autoref{algoritmo:mce-k} for MCE-$k$. 
Firstly, for every subset $I \subset \{1, \dots, m\}$, such that $|I| = k$, the algorithm for MCE is invoked for the instance $(\Pp, \Ww, \{(a_j, b_j) : j \in I\})$; that is, an instance where only the ellipses in $I$ are present.
After that, by keeping track of the utilized ellipses' costs for every $I \subset \{1, \dots, m\}$, an optimal solution can be obtained.
This simple adjustment achieves a run-time complexity of $\bigO(\binom{m}{k} \times n^{2k}) = \bigO(n^{3m})$.

\begin{algoritmo}
    \caption{Algorithm for MCE}\label{algoritmo:mce}

    \begin{algorithmic}[1]
        \Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, a list of weights $\Ww=\{w_1, \dots, w_n\}$, and a list of shape parameters $\Rr=\{(a_1, b_1), \dots, (a_m, b_m)\}$.}

        \Ensure{An optimal solution for MCE.}

        \item[]
        \Procedure{$MCE$}{$\Pp, \Ww, \Rr$}
	    \State \Return $MCE_{bt}(\Pp, \Ww, \Rr, 1)$
        \EndProcedure

        \Procedure{$MCE_{bt}$}{$Z, \Ww, \Rr, j=1$}
        \If{$j = m+1$}
        \State \Return $0$
        \EndIf

        \State $(q_j^*, \dots, q_m^*) \gets (0, \dots, 0)$

        \State $S_j \gets \textnormal{CLS-MCE}(Z, a_j, b_j)$
        \For{$q_j \in S_j$}
        \State $Cov \gets \Pp \cap E_j(q_j)$
        \State $(q_{j+1}, \dots, q_m) \gets MCE_{bt}(Z \setminus Cov, \Ww, \Rr, j+1)\}$

        \If{$w(Cov) + w(\cup_{k=j+1}^m Z \cap E_k(q_k)) >  w(\cup_{k=j}^m Z \cap E_k(q_k^*))$}
        \State $(q_j^*, \dots, q_j^*) \gets(q_j, \dots, q_m)$
        \EndIf
        \EndFor

        \State \Return $(q_j^*, \dots, q_m^*)$
        \EndProcedure
    \end{algorithmic}
\end{algoritmo}

\begin{algoritmo}
	\caption{Algorithm for MCE-$k$}\label{algoritmo:mce-k}

	\begin{algorithmic}[1]
   \Require{A set of points $\Pp=\{p_1,\dots,p_n\}$, a list of weights $\Ww=\{w_1, \dots, w_n\}$, a list of shape parameters $\Rr=\{(a_1, b_1), \dots, (a_m, b_m)\}$, a list of costs $\Cc=\{c_1, \dots, c_m\}$, and $k\in \mathbb{N}$.}
	\Ensure{An optimal solution for MCE-$k$.}

@@

\Procedure{MCE-$k$}{$\Pp, \Ww, \Rr, \Cc, k$}
	\State $I^* = \{i_1^*, \dots, i_k^*\}\gets \{1, \dots, k\}$
	\State $Q^* = (q_1^*, \dots, q_k^*) \gets (0, \dots, 0)$

	\ForAll{$I=\{i_1, \dots, i_k\} \subset \{1, \dots, m\}$, such that $|I|=k$}

		\State $(q_1, \dots, q_k) \gets MCE(\Pp, \Ww, \{(a_j, b_j) \in \Rr: j \in I\})$

		\If{$w(\bigcup_{j=1}^k \Pp \cap E_{i_j}(q_j)) - \sum_{j\in I} c_j > w(\bigcup_{j=1}^k \Pp \cap E_{i_j^*}(q_j^*)) - \sum_{j\in I^*}c_{j}$}
			\State $Q^* \gets (q_1, \dots, q_k)$
			\State $I^* \gets I$
		\EndIf
	\EndFor

	\State \Return $I^*, Q^*$
\EndProcedure
	\end{algorithmic}
\end{algoritmo}
